title: "Danish Kings"
author: "Sofia Maria Lerche Jones"
date: "2025-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The task here is to load your Danish Monarchs csv into R using the `tidyverse` toolkit, calculate and explore the kings' duration of reign with pipes `%>%` in `dplyr` and plot it over time.

## Load the kings

Make sure to first create an `.Rproj` workspace with a `data/` folder where you place either your own dataset or the provided `kings.csv` dataset. 

1. Look at the dataset that are you loading and check what its columns are separated by? (hint: open it in a plain text editor to see)

List what is the

separator: `,`  

2. Create a `kings` object in R with the different functions below and inspect the different outputs.
-  `read.csv()`
-  `read_csv()`
-  `read.csv2()` 
-  `read_csv2()`
```{r eval=TRUE}
# Load dataset using different functions
kings1 <- read.csv("data/kings.csv")
kings2 <- read_csv("data/kings.csv", na = "NULL")
kings3 <- read.csv2("data/kings.csv")
kings4 <- read_csv2("data/kings.csv")
```

Answer: 
1. Which of these functions is a `tidyverse` function?  
   - `read_csv()` is a tidyverse function.  

2. What is the result of running `class()` on the `kings` object created with a tidyverse function?  
   - `class(kings2)` returns `"spec_tbl_df" "tbl_df" "tbl" "data.frame"`.  

3. How many columns does the object have when created with these different functions?  
   ```{r}
   glimpse(kings2)  # Shows 11 columns
   ```

4. Show the dataset so that we can see how R interprets each column  
   ```{r}
   head(kings2)
   glimpse(kings2)
   ```
   - R interprets all observations as "characters," meaning it does not recognize years as numbers.  

```{r eval = TRUE}
# Assign kings2 as the working dataset
kings <- kings2

class(kings)

head(kings)

glimpse(kings)
```

## Calculate the duration of reign for all the kings in your table

```{r}
# Handle missing values
kings <- na.omit(kings)

# Create duration column
kings_mid_year <- kings %>% 
  mutate(Start_year = as.numeric(Start_year),
         End_year = as.numeric(End_year),
         duration = End_year - Start_year,
         mid_year = Start_year + duration / 2) %>% 
  glimpse()
```

## Calculate the average duration of reign for all rulers

```{r}
# Calculate mean using base R
avg_duration <- mean(kings_mid_year$duration, na.rm = TRUE)

# Calculate mean using tidyverse
kings_mid_year %>%
  summarise(avg_duration = mean(duration, na.rm = TRUE))
```

## How many and which kings enjoyed a longer-than-average duration of reign?

```{r}
# Filter kings who ruled longer than average
long_reigning_kings <- kings_mid_year %>%
  filter(duration > avg_duration) %>%
  select(Name, duration) %>% 
  print(n = Inf)
```

## How many days did the three longest-ruling monarchs rule?

```{r}
# Find longest-ruling kings and calculate total ruling days
kings_mid_year %>% 
  arrange(desc(duration)) %>% 
  slice(1:3) %>% 
  mutate(Days = duration * 365 + floor(duration / 4)) %>% 
  select(Name, duration, Days)
```

## Challenge: Plot the kings' duration of reign through time

```{r}
# Create plot
ggplot(kings_mid_year, aes(x = mid_year, y = duration)) +
  geom_point() +
  geom_smooth() +
  labs(
    title = "Duration of Danish Kings' Reigns Over Time",
    x = "Mid-Year of Reign",
    y = "Duration (Years)"
  ) +
  theme_bw() +
  theme(text = element_text(size = 14))
```

And to submit this rmarkdown, knit it into html. But first, clean up the code chunks, adjust the date, rename the author and change the `eval=FALSE` flag to `eval=TRUE` so your script actually generates an output.  
Well done!